---
title: "Statistic course Murray 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Git 

Git is a distributed version control system
Git stores snapshots of the filesystem 
The system consists of three trees:

  1. Workspace = filesystem
      - untracked --> uncommitted stages (junk) untracked, git does not know about it
  2. Staging Area = what to include in next snapshot what to commit
      -file is staged, git knows about it but it is not committed to snapshot
3. Local Repository = saved
      -savely commited


Files can be in one of four states:
1. untracked
2. staged
3. committed
4. modified

##Git configuration

You only have to do it after new installation.
Type this into the shell (under small cogwheel --> shell)

```{r}
#git config --global user.name "Your name"
#git config --global user.email "Your emailadress"
```
- Restart R studio after configuration
- check under tools --> global options --> git --> check if R studio can find git
- The name and email adress does not need to be the same name than your git hub account, since it is not linked. 
- The purpose is to identify you so git could contact you

#R studio 

##Create a new project
- File --> new project --> empty project --> where you type in the name of the new project you can tick a box to create a git repository for this project
- You get a new tap with Git next to Environment, History, Connections, Git


### Create new project with the command line
- open cogwheel and select shell

```{r}
mkdir project_name
cd project_name
git init
#create folder with mkdir and change directory with cd to write git init
```
- go to projects in upper right corner and create in new project from existing repository

### Git - adding content
Stage the changes (add)

```{r}
git add <file(s)>
#For example:
#git add analysis.R
```

##R project folder
git creates 2 files:
- bla.Rproj --> do not need to commit it
-.blabla.gitignore 

###.gitignore

- You can write files that you don't want to upload to git into .gitignore, so git ignores them and you will never upload them
-.gitignore starts with a . since it is a hidden file
- files that you never want to push to github are data files

#command line in terminal
ls -la

gives you .git file that makes it a git repository --> if you delete it it is no longer a git repository

#Git tab:
Check status of files 
questionmarks = file exists but git does not know it
tick the box so they are added
to commit click commit and write a message with what you have done

##Command line

git commit -m 'initial commit message'

commited files are save even if you delete everything

Github used hash (40 digits) to see changes if hases are the same, the files are the same


# - History
SHA hash code include identiy of auther
if you want to go back to a chertain version you have to use the has code

by default we are in master branch but we can also make a branch if we collaborate with somebody
head = to identify where to add new commit???

if you add something to a file and save it then you can see it under git with a M

commit often since commits are cheap

after 30 days they saves only the differences not the whole file

tag: applying tags can be useful if you want to go back to a version 
add tag by klicking on zahnrad (kogg) --> shell and type in:
git tag -a 'V.1' -m 'Version 1'

in the history, a small tag with the name appears

## go back to special version:

###1. checkout
tipe in shell:
git checkout hash#

you find the hash number in the history(clock)
detached head line

you can go back to last created version using the tag

git checkout tag

###2. reset 
--> removes last node but not good if you collaborate
only if you want to delete eg data that you don't want to commit
###3. revert 
--> what did i do but you can revert it. the wrong state still exists in case others are working on it
3rd way is the best

you can only reverse 1 step at a time 
you don't need to know the number becasue it is head

##Branching 

by default we are on the main branch 

shell:
git checkout -b <Name>

##Merge braches

problem did not merge


#connect new repository from github to R studio project folder
go to git hub add new repository and 
type in shell:

git remote add origin git@github.com:Octopus-monaris/Test1.git # origin is local to 
git push -u origin master
-->push local to master branch


#Collaborate:
new project from existing project get  clone repository link

before they can push new versions you have to invite them
github repository setting collaborations send invitation to person

if you are working together on a script you have to pull the latest version

#Installing packages

install.packages('package_name')

##Check were new packages will be saved
.libPaths()

#Output:
if you only have one place than all your packages will be compatible

if you have packages in multiple locations it gets super confusing --> you get out of sinc

##Cheat sheet
www.rstudio.com/resources/cheatsheets/raw

#Rmarkdown

tool pandoc
converts md file into approx. 80 other filetypes
-integrated into rstudio

Metadata are defined in the first block

starts with
---

format is in YAML 

html is standalone one file and all images are embedded into 1 file no internet access

Rmarkdown cheat sheet
# Section 1 {#Section1}
Section heading including tag

citations:

write citation(Package name) into console

to knit into PDFs you need to install LATEX 

#Basic R
Data are stored in different objects
e.g
vector e.g this vector from 1 to 10
```{r}
1:10
```
 is the index to 
```{r}
RESP1=1:4 # does not print it
(RESP1=1:4)
#assign and prints result in one step
```
```{r}
RESP1=1:4
RESP1<-1:4
```

equal is also a way to specify an argument, so you get an error if you want to assign it in a summary (a<-lm(1:10~rnorm(10)))

vectors can store words

```{r}
RESP2 = c("big","big", "small","small")
```

```{r}
RESP3 = rnorm(4,10,3)
RESP3[1]
```

Matrix has lenght and width ( columns and rows)
matrix all columns need to be of the same type

```{r}
MAT1=cbind(RESP1, RESP3)
MAT1

```

if you want ach column different types then you need a dataframe

```{r}
DF=data.frame(RESP1,RESP2,RESP3)
DF

```

#list
anything can go into a list
used for output
for Baesian analysis
```{r}
LIST=list()
```

#tibble
dataframe but saver:
1. tibble garanties that it will fit on your screen and excludes some rows to keep it neat
2. dataframe recycle around --> in tibble will not recycle
3. tibbles can embedd lists and other dataframes into one cell

```{r}
seq(from=1, to=10, by=2)
```

Functions in R:
- you can move arguments around 
- you don't need write argument names but it is less easy to read for unfamiliear people

#get help

? function_name

there are different types of functions and if you need help you need to now exactly how they are called
overloeading of functions

write helps only in console not in Rmarkdown 

#Packages
like a book

install package = buy book
load package = put book on desk --> library(package_name)
in conflicts warning be aware that some books cover other books on desk
detach = brings book back to shelf from desk

by indicating namespace we can specify which function of which package we want to use
stats::filter

conflicts are indicated by loading packages

the function select always need namespache

dplyr::select

some people write namespace before each function to make code more robust

add block with all packages loading in the top

```{r}
ls()

```

lists what we created

when we close rstudio do you want to save workspace? choose no!!! becasue if old varables are loaded automatically, you can work with old versions without knowing it

save(RESP1, file='RESP1.RData')

# Getting data into R

```{r}
library(tidyverse)
```

Data import cheat sheet

make sure you have loaded tidyverse

always write variables in capital letters so you don't overwrite functions



#Create dataset and export it

always write variables in capital letters so you don't overwrite functions
write_csv is newer compared to write.csv
```{r}
TREATMENT=gl(2,3,6, labels=c('High','Low'))
TREATMENT
DATE = gl(3,1,6,labels =c('29/02/2000','21/02/2002','20/02/2004'))
DATE
RESPONSE=round(rnorm(6,10,3),2)
RESPONSE
test_data=data.frame(TREATMENT,DATE,RESPONSE)
test_data
write_csv(test_data, path='test_data.csv')

```

#Import dataset
best to transform spread sheets into csv and load it into R
```{r}
test_data1 = read_csv(file='test_data.csv')
```
we need to modyfy the specifications

```{r}
test_data = read_csv('test_data.csv', col_types=cols(TREATMENT=col_factor(NULL),DATE=col_date(format='%d/%m/%Y')))
```
NULL uses abc
Y = 4 digid year
y = 2 digid year

```{r}
test_data
```

tibble shows the type of the data

```{r}
knitr::kable(test_data)
```

```{r}
glimpse(test_data)
```

shows all columns and a few rows

```{r}
str(test_data)
```
lot more information


###summary

```{r}
summary(test_data)
```
#short cuts 

Ctrl + Alt + i --> insert code chunk
Ctrl + shift + m --> pipe
#Pipes
%/%
```{r}
test_data %>% summary
```
#Important data manipulation tools

arrange ()

gather
spread
unite
separate

check out cheat sheet

#Download data set from website

```{r}
load(url("http://www.flutterbys.com.au/stats/downloads/data/manipulationDatasets.RData"))

```

```{r}
ls()

```

```{r}
head(data.1)
```
#Sorting data
sort data by latitude
```{r}
arrange (data.1, LAT)
```
pipe works with any function when first thing is a dataset
```{r}
data.1a = data.1 %>% arrange(LAT)
```
sort from lowest to highest latitude

if you want to have decrease order use 

```{r}
data.1b = data.1 %>% arrange(-LAT)
```

sort by 2 columns
```{r}
data.1c = data.1 %>% arrange(Cond, Temp)
```

you can also sort by calculations
```{r}
data.1d = data.1 %>% arrange(Lat+Temp)
```

sort by between and condition
```{r}
data.1d = data.1 %>% arrange(Between, Cond)
```

```{r}
data.1e = data.1 %>% arrange(Cond, Temp/LAT)
```


#Categorical variables
```{r}
data.1$Cond
```

```{r}
data.1[,'Cond']
```
out of level due to alphabetical order 

routine is order level and names directly after reading it in

only changes levels but will not arrange data
```{r}
data.3<- data.1 %>% mutate(Cond=recode_factor(Cond,'L'='Low', 'M'='Medium', 'H'='High'))
data.3
levels(data.3$Cond)
```

```{r}
data.1 %>% dplyr::select(Between,Plot,Temp,Cond,Time) %>% head
```
output will be same order than in the selection function

# exclude columns
```{r}
data.1 %>% dplyr::select(-LAT, -LONG) %>% head
```

```{r}
data.1 %>% dplyr::select(Between, Plot) %>% distinct
```

is distinct the same than unique?
a lot of overlap but some differences 
distinct is written in tidyverse so it always works but unique can have troubble

#pattern matching
check out cheat sheet for more helper functions
contains()

```{r}
dplyr::select(data.1, Between, contains('L'))
```


```{r}
dplyr::select(data.1, matches('^T[a-z]m.'))
```

^= first charachter
. any characterer
[a-z] = any lower character 
3 character must be an m
* any number of
$ last character

regex there is a cheat sheet

```{r}
dplyr::select(data.1, Between:Temp)
```

this can be really dangerous since Between:Time depends on order which can change quickly


```{r}
nasa %>% dplyr::select(lat,long, starts_with('cloud')) %>% head
```

```{r}
nasa %>% dplyr::select(lat,long, matches('^cloud.*')) %>% head
```
show rep, time and other coloums that does not include pora in the name
```{r}
tikus %>% dplyr::select(rep,time, everything(), -contains('pora')) %>% head
```

